---
title: "It’s Not About the Destination but the Journey"
author: "Juliane Donahue Bombosch, Cai Ytsma, Sofia Salcedo"
date: "15 April 2018"
output: 
  html_document:
    code_folding: hide
    df_print: paged
    css: mystyle.css
---

It's not about the destination but the journey. The MacLeish field station is a valuable asset for the Smith community. It has approximately four miles of maintained trails accessible to visitors. Ensuring accessibility is important for all levels of hiking experience and will help curate a fun experience for visitors. We aimed to classify these trails as “Easy,” “Moderate,” or “Difficult” based on their length and change in elevation. We will utilize spatial data to develop a heuristic that will classify these (or any other) trails based on their difficulty.

The data set used is the macleish package authored by Ben Baumer, as well as other contributors. This data set contains map information of the 260-acre campus, as well as live weather data taken at 10 minute intervals and stored on a computer in McConnell Hall at Smith College. 

In this analysis, we used the maps layer, focusing on mapped points of interest, elevation, trails to map the difficulty of the trails.

```{r, message = FALSE, warning = FALSE, label = "Set Up", include = FALSE, error = FALSE}
library(tidyverse)
library(dplyr)
library(sf)
library(etl)
library(macleish)
library(leaflet)

devtools::install_github("beanumber/macleish")
```

```{r, message = FALSE, warning = FALSE, include = FALSE, error = FALSE}
elev <- macleish_layers[["contours_3m"]] %>%
   st_transform(4326) 

 trails_inbound <- elev %>%
   st_intersection(macleish_layers[["trails"]])
 
 trails_pt<- st_cast(trails_inbound, "POINT")

trails <- macleish_layers[["trails"]] %>%
  group_by(name) %>%
  mutate(length = st_length(geometry)) %>%
  summarize(length_m = sum(length)/1609.34)
```

One way to distinguish the difficulty of the trails is through slope. Below are tables breaking up the difficulty into "easy," "moderate," and "hard." These tend to be arbitrary classifications, but are necessary for to distinguish a more intensive hike, such as the Vernal Pool Loop compared to the walk into the station at the beginning of the driveway.

```{r, results = 'asis', message = FALSE, warning = FALSE, label = "Difficulty Level", error = FALSE}
trails_slope <- data.frame(trails_inbound) %>%
  group_by(name) %>%
  mutate(slope = abs(c(NA, diff(ELEV_M))), len = abs(c(NA, diff(SHAPE_LEN)))) %>%
  filter(name != "Easy Out", !is.na(slope)) %>%
  select(name, SHAPE_LEN, ELEV_M, slope, len) %>%
  summarize(total_len = sum(SHAPE_LEN), tot_change = sum(slope), avg_slope_m = tot_change/total_len)

x <- 10/max(trails_slope$avg_slope_m)
y <- 10/max(trails_slope$total_len)

trails_slope <- trails_slope %>%
  mutate(slope_level = avg_slope_m * x, len_level = total_len * y)

```

```{r}
writeLines("td, th { padding : 6px } th { background-color : blue ; color : white; border : 1px solid white; } td { color : black ; border : 1px solid blue }", con = "mystyle.css")
  knitr::kable(trails_slope, format = "html") 
```

```{r, message = FALSE, warning = FALSE, label = "Our Map Code", error = FALSE}
# other variables to be mapped
buildings<- macleish_layers[["boundary"]] %>%
  st_intersection(macleish_layers[["buildings"]])
landmarks<-  macleish_layers[["boundary"]] %>%
  st_intersection(macleish_layers[["landmarks"]])
water<-  macleish_layers[["boundary"]] %>%
  st_intersection(macleish_layers[["streams"]])
meepmeep <- colorFactor(topo.colors(9), trails$name, 
                       palette = c('#2b1344','#730be5','#ff2100','#ff907f','#7ffff4', '#f7e165','#777568','#ff0083','#FFBBFF'))

campsite <- tribble(
  ~point, ~lon, ~lat,
  "Group Campsite", -72.678154, 42.450976, 
  "Remote campsite", -72.679581, 42.458549
)
camp_neu <- campsite %>%
  st_as_sf(coords = c("lon", "lat"))


redmarkercustom <- makeIcon(
  iconUrl = "http://chart.apis.google.com/chart?chst=d_map_pin_letter&chld=%E2%80%A2|e85141&chf=a,s,ee00FFFF",
  iconWidth = 8, iconHeight = 12,
  iconAnchorX = 10, iconAnchorY = 10
)

campsite_marker <- makeIcon(
  iconUrl = "https://upload.wikimedia.org/wikipedia/commons/9/9a/Legenda_pole_namiotowe.svg",
  iconWidth = 10, iconHeight = 12,
  iconAnchorX = 10, iconAnchorY = 10
)

Our_map <- leaflet(data = c( trails_inbound)) %>%
  addTiles(group = "OpenStreetMap") %>%
  addProviderTiles("Stamen.TonerLite", group = "Toner Lite") %>% 
  addPolygons(data = macleish_layers[["boundary"]], weight = 3, color = "#01561d", fillOpacity = 0.5)%>%
  addPolylines(data = macleish_layers[["boundary"]], weight = 3, color = "#185b2e", opacity = 0.65) %>%
  addCircleMarkers(data = st_cast(trails_inbound, "POINT"), color = "purple",radius = 2, popup = ~paste("Trail:",name), group = "Trail Segments") %>%
 addPolylines(data = trails, weight = 2, popup = ~name, color = ~meepmeep(trails$name)) %>%# here is where we would start adding our trails by difficulty average. :)
  addPolylines(data = elev, weight = 0.78, color = "#004400")%>%
  addPolygons(data = buildings, color = "202120", stroke = FALSE, fillOpacity = 1, weight = 1, group = "Buildings and Landmarks")%>%
  addMarkers(data = landmarks, 
             popup = ~Label, group = "Buildings and Landmarks", icon = redmarkercustom )%>%
  addMarkers(data = camp_neu, 
             popup = ~point, group = "Buildings and Landmarks", icon = campsite_marker )%>%
  addLayersControl(
    overlayGroups = c( "Buildings and Landmarks", "Trail Segments"), # we will add polylines for the level of difficulty then we can group by category.
    options = layersControlOptions(collapsed = FALSE))


Our_map


```
Length is more easily understood and can just be visually understood in the map above. The Snowmobile Trail is the longest, at 1.6 miles and would be classified as "moderate" based off this, while the entrance and exit trails would be considerably less mileage and rank at an easier level.

There is a very clear use for the ranking of difficulty: a map can be made to be displayed at Macleish that shows both the difficulty by trail length as well as average slope over the hike. With this information, visitors will be more empowered to choose their trail, rather than attempting to figure out a difficulty by the name. This would be a low cost solution: requiring a board at the beginning of the trails, as well as color coded blazes on trees throughout the trails. A good time to mark trails and create a map at the beginning would be during the early summer, when there are interns available at the station who are engaged in similar projects. Blazing could occur concurrently with trail maintenance to further streamline the process. This would require paint and general maintanence for future years, but nothing too labor intensive for the outcome. 

The Pioneer Valley is a beautiful location, and one of the best ways to experience the valley is through hiking. Many Smith College students want to be able to access this, but most hiking trails are difficult to reach on the Pioneer Valley Public Transport. For example, it takes about thirty minutes to go to Fitzgerald Lake on the bus, but in terms of elevation change and length, the conservation area is far from exciting. The access to a range of trail difficulty is a valuable asset to the students and part of that access is providing information about the difficulty level of different trails both through the elevation changes, as well as length.

Footnotes: 
[^1] Link to the [repository](https://github.com/jbombosch/traildifficult)
[^2] Link to the [macleish repository](https://github.com/beanumber/macleish)